Informing policy via dynamic models: Cholera in Haiti

### 2023-08-22 ms-plos.Rnw
Moved the figures and tables, made figure titles

%%%% parameters %%%%%%%%%%%
\newcommand\Wsat{W_{\mathrm{sat}}}
\newcommand\muIR{\mu_{IR}}
\newcommand\muEI{\mu_{EI}}
\newcommand\transmission{\beta}
\newcommand\seasAmplitude{a}
\newcommand\rainfallExponent{r}
\newcommand\muRS{\mu_{RS}}
\newcommand\vaccineEfficacy{\vartheta}
\newcommand\muBirth{\mu_S}
\newcommand\muDeath{\delta}
\newcommand\choleraDeath{\delta_{C}}
\newcommand\symptomFrac{f}
\newcommand\asymptomRelativeInfect{\epsilon}
\newcommand\asymptomRelativeShed{\epsilon_{W}}
\newcommand\Wbeta[1]{\beta_{W#1}}
\newcommand\Whur[1]{\beta_{W#1}^{hm}}
\newcommand\hHur[1]{h_{#1}^{hm}}
\newcommand\tHur{t_{hm}}
\newcommand\Wremoval{\delta_W}
\newcommand\Wshed{\mu_W}
\newcommand\mixExponent{\nu}
\newcommand\sigmaProc{\sigma_{\mathrm{proc}}}
\newcommand\reportRate{\rho}
\newcommand\obsOverdispersion{\psi}
\newcommand\phaseParm{\phi}
\newcommand\transmissionTrend{\zeta}
\newcommand\Binit{\xi}
\newcommand\vaccClass{Z}
\newcommand\vaccCounter{z}
\newcommand\modelCounter{m}
\newcommand\missing{---}
\newcommand\fixed{\color{blue}}
\newcommand\demography{\bullet}
\newcommand\code[1]{\texttt{#1}}
\newcommand\paramVec{\theta}
\newcommand\childReduce{q}

@@

# Plot reported cholera cases, by department. Creates Figure 1.
ggplot(dep_plot_df, aes(x = date, y = cases + 1)) +
  facet_wrap(~dep, nrow = 2, labeller = dep_labeller) +
  geom_line() +
  theme(axis.title.x = element_blank()) +
  ylab('Reported Cases') +
  scale_y_log10(
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  scale_x_date(
    date_labels = "'%y",
    breaks = seq.Date(
      from = as.Date("2011-01-01"),
      to = as.Date("2019-01-01"),
      by = '2 years'
    )
  )
@
\caption{\label{fig:CholeraData}
\small
Weekly reported cholera cases in Haiti from October 2010 to January 2019.
}
\end{figure}

@@

The four mechanistic models of Lee et al.~\cite{lee20} were deliberately developed with limited coordination.
This allows us to treat the models as fairly independently developed expert approaches to understanding cholera transmission.
However, it led to differences in notation, and in subsets of the data chosen for analysis, that hinder direct comparison.
Here, we have put all three models into a common notational framework.
Translations back to the original notation of Lee et al.~\cite{lee20} are given in Table~S-1 of the supplement.

@@

  SEARCH2 = list(
    TOP_N = 12,
    NBPF = 100,
    NP = 2500,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 2500,
    NREPS_EVAL = 18,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = FALSE,
    KEEP_LIK_MAT = TRUE
  )

  # SEARCH3 = list(  # Not used, but was used for some preliminary results.
  #   TOP_N = 9,
  #   NBPF = 100,
  #   NP = 2000,
  #   SPAT_REGRESSION = 0.05,
  #   NREPS = 8,
  #   NP_EVAL = 2000,
  #   NREPS_EVAL = 18,
  #   RW_SD = rw_3.sd,
  #   COOLING = 0.5,
  #   KEEP_TRACES = FALSE,
  #   KEEP_LIK_MAT = TRUE
  # )
}

@@

In this article we calibrate the parameters of each of our models by maximizing the likelihood, as described in Eq.~\myeqref{eq:likedef}.
In the following subsections we describe in detail our approach to calibrating the three proposed mechanistic models to observed cholera incidence data.

\begin{figure}[!h]
\centering
<<Beta_trend_Figure, fig.height=2.4, fig.width=3.6, fig.align='center'>>=
ggplot() +
  geom_point(data = h1_prof_res, aes(x = betat, y = logLik)) +
  geom_line(data = mcap_results$fit, aes(x = parameter, y = smoothed), col = 'blue') +
  geom_vline(xintercept = mcap_results$ci[1], linetype = 'dashed') +
  geom_vline(xintercept = mcap_results$ci[2], linetype = 'dashed') +
  geom_vline(xintercept = mcap_results$mle, col = 'blue') +
  labs(x = "Linear Trend in Transmission", y = 'Log Likelihood') +
  theme(axis.text = element_text(size = 8),
        axis.title = element_text(size = 10))
@
\caption{\label{fig:betat}
Monte Carlo adjusted profile of $\transmissionTrend$.
The blue curve is the profile, the blue line indicates the MLE, and the dashed lines indicate the $95\%$ confidence interval.
}
\end{figure}

# Plot the data, median value of the simulation, and estimated quantiles.
ggplot() +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  geom_line(data = quants, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_ribbon(  # Plot confidence region
    data = quants,
    aes(x = date, ymin = q025 + 1, ymax = q975 + 1),
    alpha = 0.5
  ) +
  theme(
    axis.title.x = element_blank(),
    axis.text = element_text(size = 8),
    axis.title.y = element_text(size = 10)
  )+
  ylab('Reported cholera cases') +
  geom_vline(  # Show change from epidemic -> endemic models.
    xintercept = lubridate::weeks(232) + lubridate::ymd("2010-10-16"),
    linetype = 'dashed'
  ) +
  scale_y_log10(  # Plot on the log-scale
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  scale_x_date(  # Format the date on x-axis.
    date_labels = "'%y",
    breaks = seq.Date(
      from = as.Date("2011-01-01"),
      as.Date("2019-01-01"),
      by = '1 year'
    )
  )
@
\caption{\label{fig:mod1fit}
Simulations from Model~1 compared to reported cholera cases.
The black curve is observed data, the blue curve is median of 500 simulations from the fitted model, and the vertical dashed line represents break-point when parameters are refit.
}
\end{figure}

\begin{figure}[!h]
<<Mod2Fit_and_Scenarios_Figure, fig.height=1.9, fig.width=4.8, fig.align='center'>>=
ggplot() +
  geom_line(data = h2_traj, aes(x = date, y = Ctotal + 1), col = 'blue') +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  geom_line(data = filter(mod2_VaccScenarios$all_trajs,
                          year >= max(h2_traj$year)),
            aes(x = as.Date(date), y = ReportedCases + 1, color = scenario)) +
  theme(axis.title.x = element_blank(),
        legend.title = element_blank(),
        axis.title.y = element_text(size = 9),
        axis.text = element_text(size = 8)) +
  ylab("Reported cholera cases") +
  scale_color_manual(values = c("V0" = '#f4a582',
                                "V1" = '#d6604d',
                                "V2" = '#b2182b',
                                "V3" = '#92c5de',
                                "V4" = '#4393c3')) +
  # geom_vline(xintercept = yearsToDate(2014.161), linetype = 'dashed') +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_x_date(date_labels = "'%y", breaks = seq.Date(from = as.Date("2010-01-01"), as.Date("2030-01-01"), by = '2 years'))
@
\caption{\label{fig:mod2Traj}
Simulated trajectory of Model~2 (blue curve) and projections under the various vaccination scenarios.
Reported cholera incidence is shown in black.}
\end{figure}

# Plot the reported cases and confidence region for each department.
ggplot() +
  geom_line(data = h3_quants, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_line(data = dep_plot_df, aes(x = date, y = cases + 1), col = 'black') +
  geom_ribbon(  # plot confidence region.
    data = h3_quants,
    aes(x = date, ymin = q05 + 1, ymax = q95 + 1),
    alpha = 0.4
  ) +
  facet_wrap(
    ~factor(dep, levels = plot_order), nrow = 2, labeller = dep_labeller
  ) +
  labs(y = 'Log Reported Cholera Cases') +
  scale_x_date(  # Format the date on the X-axis.
    date_labels = "'%y",
    breaks = seq.Date(
      from = as.Date("2011-01-01"),
      to = as.Date("2019-01-01"),
      by = '2 years'
    )
  ) +
  theme(axis.title.x = element_blank()) +
  scale_y_log10(  # Plot on log-scale
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  )
@
\caption{\label{fig:h3spatsims}
Simulations from initial conditions using the spatially coupled version of Model~3.
The black curve represents true case count, the blue line the median of 500 simulations from the model, and the gray ribbons representing $95\%$ confidence interval.
}
\end{figure}

<<Table 2 Input, echo=FALSE, eval=TRUE>>=
NREPS_EVAL <- switch(RUN_LEVEL,   3,   8,    36)
NP_EVAL    <- switch(RUN_LEVEL, 100, 500,  5000)

@@

<<CalculateARMA, echo=FALSE, message=FALSE, include=FALSE, cache=TRUE>>=

@@

<<Table 2 Lee Input, echo=FALSE, include=FALSE>>=
mod1_lee_n_params <- 20
mod1_lee_aic <- 2 * mod1_lee_n_params - 2 * mod1_lee_ll

mod2_lee_aic <- 2 * h2_fit$leeFit_n_params - 2 * h2_fit$leeFit_ll
@

<<Lee3 Sub Calculations, echo=FALSE, include=FALSE, message=FALSE>>=
lee3_NP   <- switch(RUN_LEVEL, 50, 250, 5000)
lee3_NREP <- switch(RUN_LEVEL,  3,   8,   36)

@@

\begin{table}
\centering

\begin{tabular}{|c|c|c|c|}
\hline
 & \thead{Model~1} & \thead{Model~2} & \thead{Model~3}
\\
\hline
\hline
\multirow{2}{*}{Log-likelihood} &
  $\Sexpr{myround(mod1_ll, 1)}$ &
  $\Sexpr{myround(mod2_ll, 1)}$ &
  $\Sexpr{myround(mod3_ll, 1)}$ \\
    & ($\Sexpr{myround(mod1_lee_ll, 1)}$)\footnotemark[1] &
  ($\Sexpr{myround(h2_fit$leeFit_ll, 1)}$) &
  ($\Sexpr{myround(lee3_results[["all_ll_eval"]], 1)}$)\footnotemark[2]
\\
\hline
Number of &
  $\Sexpr{as.character(mod1_n_params)}$ &
  $\Sexpr{as.character(mod2_n_params)}$ &
  $\Sexpr{as.character(mod3_n_params)}$ \\
 Fit Parameters &
 (\Sexpr{as.character(mod1_lee_n_params)}) &
 ($\Sexpr{as.character(h2_fit$leeFit_n_params)}$) &
 ($\Sexpr{as.character(lee3_results$n_params, 1)}$)
\\
\hline
\multirow{2}{*}{AIC} &
  $\Sexpr{myround(mod1_aic, 1)}$ &
  $\Sexpr{myround(mod2_aic, 1)}$ &
  $\Sexpr{myround(mod3_aic, 1)}$ \\
  & ($\Sexpr{myround(mod1_lee_aic, 1)}$)\footnotemark[1] &
  ($\Sexpr{myround(mod2_lee_aic, 1)}$) &
  ($\Sexpr{myround(mod3_lee_coupled_aic, 1)}$)\footnotemark[2]
\\
\hline
\thead{Log-ARMA(2,1) \\ AIC} &
  $\Sexpr{myround(-2 * ARMA_benchmarks$m1 + 2 * 5, 1)}$ &
  $\Sexpr{myround(-2 * ARMA_benchmarks$m2 + 2 * 5, 1)}$ &
  $\Sexpr{myround(-2 * ARMA_benchmarks$m3 + 2 * 5, 1)}$
\\
\hline
\end{tabular}
\caption{\label{tab:likes}AIC values for each model compared to their ARMA benchmark.
Values in parentheses are corresponding values using \cite{lee20} parameter estimates.  $\textsuperscript{1}$The reported likelihood is an upper bound of the likelihood of the \cite{lee20} model. $\textsuperscript{2}$\cite{lee20} fit Model~3 to a subset of the data (March 2014 onward, excluding data from Ouest in 2015-2016).
On this subset, their model has a likelihood of $\Sexpr{myround(lee3_results$sub_ll_eval, 1)}$.
On this same subset, our model has a likelihood of $\Sexpr{myround(mod3_eval_res$subset_ll, 1)}$.
See Sec.~S6 of the supplement material for more details on estimating the likelihood of the Lee et al.~\cite{lee20} models.
}
\end{table}

\begin{figure}[!h]
\centering
<<Model1_Seasonality_Figure, fig.height=2.9, fig.width=5, fig.align='center'>>=
cowplot::plot_grid(gg_rain, gg_trans, align = "v", ncol = 1)
@
\caption{\label{fig:h1SeasRain}
(Top) weekly rainfall in Haiti, lighter colors representing more recent years.
(Bottom) estimated seasonality in the transmission rate (dashed line) plotted alongside mean rainfall (solid line).
The outsided effect of rainfall in the fall may be due to Hurricane Matthew, which struck Haiti in October of 2016 and resulted in an increase of cholera cases in the nation.
}
\end{figure}

% Place figure captions after the first paragraph in which they are cited.
\begin{figure}[!h]
\caption{{\bf Bold the figure title.}
Figure caption text here, please use this space for the figure panel descriptions instead of using subfigure commands. A: Lorem ipsum dolor sit amet. B: Consectetur adipiscing elit.}
\label{fig1}
\end{figure}

\begin{figure}[!h]
<<Elimination_Probs_Figure, fig.height=2.2>>=
ggplot(tidyr::drop_na(all_prob_elims), aes(x = time, y = elim_prob, col = mod)) +
  geom_line() +
  facet_wrap(~scenario, nrow = 1) +
  scale_color_manual(values = c("Model 1" = "#377eb8", "Model 3" = "#e41a1c")) +
  ylab("Probability of Elimination (%)") +
  theme(axis.title.x = element_blank(),
        legend.position = 'bottom',
        legend.title = element_blank(),
        axis.text.x = element_text(angle = 35, hjust = 1),
        legend.margin=margin(c(-5, 0, -3, 0)),
        legend.box.margin = margin(c(-5, 0, -3, 0))) +
  scale_x_date(date_labels = "'%y", breaks = seq.Date(from = as.Date("2010-01-01"), as.Date("2030-01-01"), by = '2 years'))
@
\caption{\label{fig:elimProbs}
Probability of elimination across simulations for a 10 year period.
Compare to Fig.~3A of Lee et al.~\cite{lee20}.}
\end{figure}

%%% table tttttttttt

@@

>>>>

%%%% parameters %%%%%%%%%%%
\newcommand\Wsat{W_{\mathrm{sat}}}
\newcommand\muIR{\mu_{IR}}
\newcommand\muEI{\mu_{EI}}
\newcommand\transmission{\beta}
\newcommand\seasAmplitude{a}
\newcommand\rainfallExponent{r}
\newcommand\muRS{\mu_{RS}}
\newcommand\vaccineEfficacy{\vartheta}
\newcommand\muBirth{\mu_S}
\newcommand\muDeath{\delta}
\newcommand\choleraDeath{\delta_{C}}
\newcommand\symptomFrac{f}
\newcommand\asymptomRelativeInfect{\epsilon}
\newcommand\asymptomRelativeShed{\epsilon_{W}}
\newcommand\Wbeta[1]{\beta_{W#1}}
\newcommand\Whur[1]{\beta_{W#1}^{hm}}
\newcommand\hHur[1]{h_{#1}^{hm}}
\newcommand\tHur{t_{hm}}
\newcommand\Wremoval{\delta_W}
\newcommand\Wshed{\mu_W}
\newcommand\mixExponent{\nu}
\newcommand\sigmaProc{\sigma_{\mathrm{proc}}}
\newcommand\reportRate{\rho}
\newcommand\obsOverdispersion{\psi}
\newcommand\phaseParm{\phi}
\newcommand\transmissionTrend{\zeta}
\newcommand\Binit{\xi}
\newcommand\vaccClass{Z}
\newcommand\vaccCounter{z}
\newcommand\modelCounter{m}
\newcommand\missing{---}
\newcommand\fixed{\color{blue}}
\newcommand\demography{\bullet}
\newcommand\code[1]{\texttt{#1}}
\newcommand\figTitle{\bf}
\newcommand\paramVec{\theta}
\newcommand\childReduce{q}

@@

# Plot reported cholera cases, by department. Creates Figure 1.
ggplot(dep_plot_df, aes(x = date, y = cases + 1)) +
  facet_wrap(~dep, nrow = 2, labeller = dep_labeller) +
  geom_line() +
  theme(axis.title.x = element_blank()) +
  ylab('Reported Cases') +
  scale_y_log10(
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  scale_x_date(
    date_labels = "'%y",
    breaks = seq.Date(
      from = as.Date("2011-01-01"),
      to = as.Date("2019-01-01"),
      by = '2 years'
    )
  )
@
\caption{\label{fig:CholeraData}
{\figTitle Weekly cholera cases.}
Weekly reported cholera cases in Haiti from October 2010 to January 2019 for each of the 10 administrative departments.
}
\end{figure}

@@

@@

  SEARCH2 = list(
    TOP_N = 12,
    NBPF = 100,
    NP = 2500,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 2500,
    NREPS_EVAL = 18,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = FALSE,
    KEEP_LIK_MAT = TRUE
  )
}

@@

<<AICtableInput, echo=FALSE, eval=TRUE>>=
NREPS_EVAL <- switch(RUN_LEVEL,   3,   8,    36)
NP_EVAL    <- switch(RUN_LEVEL, 100, 500,  5000)

@@

<<CalculateARMALikelihood, echo=FALSE, message=FALSE, include=FALSE, cache=TRUE>>=

@@

<<AIC Lee et al, echo=FALSE, include=FALSE, message=FALSE>>=
lee3_NP   <- switch(RUN_LEVEL, 50, 250, 5000)
lee3_NREP <- switch(RUN_LEVEL,  3,   8,   36)

mod1_lee_n_params <- 20
mod1_lee_aic <- 2 * mod1_lee_n_params - 2 * mod1_lee_ll

mod2_lee_aic <- 2 * h2_fit$leeFit_n_params - 2 * h2_fit$leeFit_ll

@@

The four mechanistic models of Lee et al.~\cite{lee20} were deliberately developed with limited coordination.
This allows us to treat the models as fairly independently developed expert approaches to understanding cholera transmission.
However, it led to differences in notation, and in subsets of the data chosen for analysis, that hinder direct comparison.
Here, we have created a common notational framework that facilitates model comparison, and put all comparable model parameters---including parameters that were estimated or held constant---into Table~\ref{tab:params}.
% Here, we have put all three models into a common notational framework.
Translations back to the original notation of Lee et al.~\cite{lee20} are given in Table~S-1 of the supplement. % TODO: Check S-1 table matches.

@@

In this article we calibrate the parameters of each of our models by maximizing the likelihood, as described in Eq.~\myeqref{eq:likedef}.
The likelihood for each of the fitted models---and the corresponding AIC values for model comparisons that include an adjustment for the number of calibrated parameters---is provided in Table~\ref{tab:likes}
In the following subsections we describe in detail our approach to calibrating the three proposed mechanistic models to observed cholera incidence data.

\begin{table}[!h]
\centering

\begin{tabular}{|c|c|c|c|}
\hline
 & \thead{Model~1} & \thead{Model~2} & \thead{Model~3}
\\
\hline
\hline
\multirow{2}{*}{Log-likelihood} &
  $\Sexpr{myround(mod1_ll, 1)}$ &
  $\Sexpr{myround(mod2_ll, 1)}$ &
  $\Sexpr{myround(mod3_ll, 1)}$ \\
    & ($\Sexpr{myround(mod1_lee_ll, 1)}$)\footnotemark[1] &
  ($\Sexpr{myround(h2_fit$leeFit_ll, 1)}$) &
  ($\Sexpr{myround(lee3_results[["all_ll_eval"]], 1)}$)\footnotemark[2]
\\
\hline
Number of &
  $\Sexpr{as.character(mod1_n_params)}$ &
  $\Sexpr{as.character(mod2_n_params)}$ &
  $\Sexpr{as.character(mod3_n_params)}$ \\
 Fit Parameters &
 (\Sexpr{as.character(mod1_lee_n_params)}) &
 ($\Sexpr{as.character(h2_fit$leeFit_n_params)}$) &
 ($\Sexpr{as.character(lee3_results$n_params, 1)}$)
\\
\hline
\multirow{2}{*}{AIC} &
  $\Sexpr{myround(mod1_aic, 1)}$ &
  $\Sexpr{myround(mod2_aic, 1)}$ &
  $\Sexpr{myround(mod3_aic, 1)}$ \\
  & ($\Sexpr{myround(mod1_lee_aic, 1)}$)\footnotemark[1] &
  ($\Sexpr{myround(mod2_lee_aic, 1)}$) &
  ($\Sexpr{myround(mod3_lee_coupled_aic, 1)}$)\footnotemark[2]
\\
\hline
\thead{Log-ARMA(2,1) \\ AIC} &
  $\Sexpr{myround(-2 * ARMA_benchmarks$m1 + 2 * 5, 1)}$ &
  $\Sexpr{myround(-2 * ARMA_benchmarks$m2 + 2 * 5, 1)}$ &
  $\Sexpr{myround(-2 * ARMA_benchmarks$m3 + 2 * 5, 1)}$
\\
\hline
\end{tabular}
\caption{\label{tab:likes}AIC values for each model compared to their ARMA benchmark.
Values in parentheses are corresponding values using Lee et al.~\cite{lee20} parameter estimates.  $\textsuperscript{1}$The reported likelihood is an upper bound of the likelihood of the Lee et al.~\cite{lee20} model. $\textsuperscript{2}$ Lee et al.~\cite{lee20} fit Model~3 to a subset of the data (March 2014 onward, excluding data from Ouest in 2015-2016).
On this subset, their model has a likelihood of $\Sexpr{myround(lee3_results$sub_ll_eval, 1)}$.
On this same subset, our model has a likelihood of $\Sexpr{myround(mod3_eval_res$subset_ll, 1)}$.
See Sec.~S6 of the supplement material for more details on estimating the likelihood of the Lee et al.~\cite{lee20} models.
}
\end{table}

\begin{figure}[!h]
\centering
<<Beta_trend_Figure, fig.height=2.4, fig.width=3.6, fig.align='center'>>=
ggplot() +
  geom_point(data = h1_prof_res, aes(x = betat, y = logLik)) +
  geom_line(data = mcap_results$fit, aes(x = parameter, y = smoothed), col = 'blue') +
  geom_vline(xintercept = mcap_results$ci[1], linetype = 'dashed') +
  geom_vline(xintercept = mcap_results$ci[2], linetype = 'dashed') +
  geom_vline(xintercept = mcap_results$mle, col = 'blue') +
  labs(x = "Linear Trend in Transmission", y = 'Log Likelihood') +
  theme(axis.text = element_text(size = 8),
        axis.title = element_text(size = 10))
@
\caption{\label{fig:betat}
{\figTitle Confidence inverval for linear trend in transmission.}
Monte Carlo adjusted profile of $\transmissionTrend$ for Model~1.
The blue curve is the profile, the blue line indicates the MLE, and the dashed lines indicate the $95\%$ confidence interval.
}
\end{figure}

# Plot the data, median value of the simulation, and estimated quantiles.
ggplot() +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  geom_line(data = quants, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_ribbon(  # Plot confidence region
    data = quants,
    aes(x = date, ymin = q025 + 1, ymax = q975 + 1),
    alpha = 0.5
  ) +
  theme(
    axis.title.x = element_blank(),
    axis.text = element_text(size = 8),
    axis.title.y = element_text(size = 10)
  )+
  ylab('Reported cholera cases') +
  geom_vline(  # Show change from epidemic -> endemic models.
    xintercept = lubridate::weeks(232) + lubridate::ymd("2010-10-16"),
    linetype = 'dashed'
  ) +
  scale_y_log10(  # Plot on the log-scale
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  ) +
  scale_x_date(  # Format the date on x-axis.
    date_labels = "'%y",
    breaks = seq.Date(
      from = as.Date("2011-01-01"),
      as.Date("2019-01-01"),
      by = '1 year'
    )
  )
@
\caption{\label{fig:mod1fit}
{\figTitle Simulations from Model~1 compared to reported cholera cases.}
The black curve is observed data, the blue curve is median of 500 simulations from the fitted model, and the vertical dashed line represents break-point when parameters are refit.
}
\end{figure}

\begin{figure}[!h]
<<Mod2Fit_and_Scenarios_Figure, fig.height=1.9, fig.width=4.8, fig.align='center'>>=
ggplot() +
  geom_line(data = h2_traj, aes(x = date, y = Ctotal + 1), col = 'blue') +
  geom_line(data = true_agg_cases, aes(x = date, y = ReportedAll + 1)) +
  geom_line(data = filter(mod2_VaccScenarios$all_trajs,
                          year >= max(h2_traj$year)),
            aes(x = as.Date(date), y = ReportedCases + 1, color = scenario)) +
  theme(axis.title.x = element_blank(),
        legend.title = element_blank(),
        axis.title.y = element_text(size = 9),
        axis.text = element_text(size = 8)) +
  ylab("Reported cholera cases") +
  scale_color_manual(values = c("V0" = '#f4a582',
                                "V1" = '#d6604d',
                                "V2" = '#b2182b',
                                "V3" = '#92c5de',
                                "V4" = '#4393c3')) +
  # geom_vline(xintercept = yearsToDate(2014.161), linetype = 'dashed') +
  scale_y_log10(labels = scales::trans_format("log10", scales::math_format(10^.x))) +
  scale_x_date(date_labels = "'%y", breaks = seq.Date(from = as.Date("2010-01-01"), as.Date("2030-01-01"), by = '2 years'))
@
\caption{\label{fig:mod2Traj}
{\figTitle Simulated trajectory of Model~2.}
The black line shows the nationally aggregated weekly cholera incidence data.
The blue curve from 2012-2019 is the trajectory of the calibrated version of Model~2.
Projections under the various vaccination scenarios, which are discussed in the \bf{Forecasts} subsection are also included.}
\end{figure}

# Plot the reported cases and confidence region for each department.
ggplot() +
  geom_line(data = h3_quants, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_line(data = dep_plot_df, aes(x = date, y = cases + 1), col = 'black') +
  geom_ribbon(  # plot confidence region.
    data = h3_quants,
    aes(x = date, ymin = q05 + 1, ymax = q95 + 1),
    alpha = 0.4
  ) +
  facet_wrap(
    ~factor(dep, levels = plot_order), nrow = 2, labeller = dep_labeller
  ) +
  labs(y = 'Log Reported Cholera Cases') +
  scale_x_date(  # Format the date on the X-axis.
    date_labels = "'%y",
    breaks = seq.Date(
      from = as.Date("2011-01-01"),
      to = as.Date("2019-01-01"),
      by = '2 years'
    )
  ) +
  theme(axis.title.x = element_blank()) +
  scale_y_log10(  # Plot on log-scale
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  )
@
\caption{\label{fig:h3spatsims}
{\figTitle Simulations from Model~3 compared to reported cholera cases.}
Simulations from initial conditions using the spatially coupled version of Model~3.
The black curve represents true case count, the blue line the median of 500 simulations from the model, and the gray ribbons representing $95\%$ confidence interval.
}
\end{figure}

\begin{figure}[!h]
\centering
<<Model1_Seasonality_Figure, fig.height=2.9, fig.width=5, fig.align='center'>>=
cowplot::plot_grid(gg_rain, gg_trans, align = "v", ncol = 1)
@
\caption{\label{fig:h1SeasRain}
{\figTitle Seasonality of Model~1 transmission compared to rainfall data.}
(Top) weekly rainfall in Haiti, lighter colors representing more recent years.
(Bottom) estimated seasonality in the transmission rate (dashed line) plotted alongside mean rainfall (solid line).
The outsided effect of rainfall in the fall may be due to Hurricane Matthew, which struck Haiti in October of 2016 and resulted in an increase of cholera cases in the nation.
}
\end{figure}

\begin{figure}[!h]
<<Elimination_Probs_Figure, fig.height=2.2>>=
ggplot(tidyr::drop_na(all_prob_elims), aes(x = time, y = elim_prob, col = mod)) +
  geom_line() +
  facet_wrap(~scenario, nrow = 1) +
  scale_color_manual(values = c("Model 1" = "#377eb8", "Model 3" = "#e41a1c")) +
  ylab("Probability of Elimination (%)") +
  theme(axis.title.x = element_blank(),
        legend.position = 'bottom',
        legend.title = element_blank(),
        axis.text.x = element_text(angle = 35, hjust = 1),
        legend.margin=margin(c(-5, 0, -3, 0)),
        legend.box.margin = margin(c(-5, 0, -3, 0))) +
  scale_x_date(date_labels = "'%y", breaks = seq.Date(from = as.Date("2010-01-01"), as.Date("2030-01-01"), by = '2 years'))
@
\caption{\label{fig:elimProbs}
{\figTitle Simulated probability of elimination using Models~1 and 3.}
Probability of cholera elimination, defined as having zero cholera infectious for at least 52 consecutive weeks, based on 10 year simulations from calibrated versions of Models~1 and 3.
Compare to Fig.~3A of Lee et al.~\cite{lee20}.}
\end{figure}
